---
title: " Adler_hmk_03"
format: gfm
editor: visual
---

# Base R and R Basics

HINT: Remember that you can get help on any function by typing `?`(function name). For instance, `?rnorm` gives help on the `rnorm()` function.

## Creating and naming variables

1.  Create a variable called `x` and use it to store the result of the calculation `(3*(4+2)`.

```{r}
x <- (3*(4+2))
```

**Question 1:** r x \^It stopped working when I tried to do it

2.  Calculate the product of `x` (from the above question) times π.

```{r}
Q2 <- x*pi
```

**Question 2:** r Q2

3.  Use the `getwd()` function to show your current working directory. Is that a good working directory, and what program do you think set it that way?

    ```{r}
    getwd()
    ```

    Question 3: Yes this is a good working directory, it is organized logically. I am not sure what program you are referring too- I set it to this directory when I saved the template file in this folder.

## Vectors

1.  Use the `c()` function to create a vector of numbers.

```{r}
numbers <- c(sample(1 : 10, size = 10, replace = F)) 
print(numbers)
```

**Question 1:** r numbers I used the sample function to create a list of 10 random numbers between 1 and 10 that do not repeat.

2.  Use the `c()` function to create a vector of characters.

```{r}
characters <- c("A", "B", "C","D","E","F")
print(characters)
```

**Question 2:** r characters

3.  Use the `:` implicit function to create a vector of integers from 1 to 10.

```{r}
integers <- c(1:10)
print(integers)
```

**Questions 3:** r integers

4.  Explain *why* the following code returns what it does. Also address whether you think this was a good decision on the part of the designers of R?

```{r, warning=FALSE}
v1 <- 1:3
v2 <- c(1:4)
v1 + v2
```

Question 4: This code is generating a vector "V1" that contains integers from 1 to 3 and it is being added to another vector "V2" that contains 1 through 4. It is adding them based on index (i.e first entry + first entry = 2) until the shortest vector is run through then it starts over from the first entry again. THis is why the last entry is 5 because 1+4 =5

5.  Explain what the following code does. It may be helpful to reference the answer to the previous question:

```{r}
c(1, 5, 9) + 3
```

Question 5: The above code is adding the integer 3 to each of the entries in the concatenated vector individually and outputing the new, unsaved vector.

6.  Remove (delete) every variable in your workspace.

```{r}
rm(list=ls())
```

Poof!

## Graphics

1.  Load the tidyverse package. **NOTE:** Be sure to use the chunk option `message=FALSE` to suppress the messages that tidyverse prints when loaded. These messages are useful in the

```{r, message=FALSE}
library(tidyverse)
message =FALSE
```

2.  Recreate the visualization of `body_mass_g` to `flipper_length_mm`, from the penguins data set, that is shown in question 8 of section 2.2.5 of [R4DS](https://r4ds.hadley.nz/data-visualize).

```{r}
library(tidyverse)
library(palmerpenguins)

glimpse(penguins)



ggplot(data=penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g, color = bill_depth_mm)) + geom_point() + geom_smooth(method ="loess")
```

Question- How do I go to a new line without messing up the function? I know the indentation is important but when I tried to do it, it was acting up.

Question - How do I determine what the default size of the points is? I was trying to change it but size= 1.5 and size = 3 looked the same, and were too large.

3.  Explain why each aesthetic is mapped at the level that it is (i.e., at the global level, in the `ggplot()` function call, or at the geom level, in the `geom_XXX()` function call). Note: A lot of different options will work, but some options are clearly better than others.

Question 3: Aesthetic is on the global level within the ggplot function because no matter how I analyze the data, the data I want to work with is not changing. In comparsion, I mapped the method of smoothing in the geom level because if I want to compare different methods or other such analyzes, I do not want my underlying data visualization to change because then I can not easily compare the results. An example would be with the method lm vs loess.
