---
title: "hmk07_badplot"
format: gfm
editor: visual
---

## The Data

MOI (Multiplicity of infection) data of Microcystis aeuroginasa infected with LM-MMA01 virus.

## Bad Plot: Reasons

-   Uses too much ink
-   Misleading for how the data in all the plots is related to each other
-   Unclear axis labels
-   No statistics shown for the data

```{r, message=FALSE}
#Packages
library(tidyverse)
library(ggplot2)
library(dplyr) 
#getwd()

#Actual version
#Load CSV
MOI_Determin<-read_csv("20231020HMK7MOI.csv")

#trying to get 2 digits for MOI in the figure legend
format(MOI_Determin$MOI,digits=2, nsmall=2)

#Figure generation: want different colors and different symbols in the legend per MOI with 2 decimal places
zero<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (0hrs)`, 
                         ))+geom_point()+labs(y="0hrs")
eight<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (8hrs)`, 
                         ))+geom_point()+labs(y="8hrs")
ten<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (10hrs)`, 
                         ))+geom_point()+labs(y="10hrs")
twelve<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (12hrs)`, 
                         ))+geom_point()+labs(y="12hrs")
fourteen<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (14hrs)`, 
                         ))+geom_point()+labs(y="14hrs")
eighteen<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (18hrs)`, 
                         ))+geom_point()+labs(y="18hrs")
twenty<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (20hrs)`, 
                         ))+geom_point()+labs(y="20hrs")
twentyfour<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (24hrs)`, 
                         ))+geom_point()+labs(y="24hrs")
thirtysiz<-ggplot(MOI_Determin, aes(x=MOI, y=MOI_Determin$`Cell Density (36hrs)`, 
                         ))+geom_point()+labs(y="36hrs")
  
library(patchwork)
my_plot<-zero+eight+ten+twelve+fourteen+eighteen+twenty+twentyfour+thirtysiz+plot_layout( ncol=3)+scale_x_discrete(guide = guide_axis(n.dodge=6))

my_plot
#ggsave("badplot.png", my_plot, width = 15, height = 10)

thirtysix<-ggplot(MOI_Determin, aes(x="", y=MOI_Determin$`Cell Density (36hrs)`, fill=MOI))+geom_bar(stat="identity", width=1 )+ coord_polar("y",start=0)

```

##Good Plot: 
- Saves ink 
- Colorblind friendly 
- Shows statistics for the data 
- Understandable for how the data is related to each other

```{r, message=FALSE}
#Packages
library(tidyverse)
library(ggplot2)
library(dplyr) 
getwd()

#Actual version
#Load CSV
df<- read_csv("2023092325_MA_MOI_Determination.csv")
glimpse(df)

#trying to get 2 digits for MOI in the figure legend
format(df$VLPMOI,digits=2, nsmall=2)

#Figure generation: want different colors and different symbols in the legend per MOI with 2 decimal places
my_good<- ggplot(df, aes(x=Time, y= Mean, color = factor(VLPMOI), shape=factor(VLPMOI))) + 
  geom_line() +
  geom_point() +
  scale_color_brewer(palette="Dark2") +
  geom_errorbar(ymin=(df$Mean - df$`Cell Density STD`), ymax=(df$Mean+ df$`Cell Density STD`), width = 0.5)+
  labs(x="Time (HRS)", y="Mean (cells/mL)",color="MOI", shape = "MOI", title=" Infection Time series: Cell Density of Microcystis") +
  theme_classic()

my_good
```
##For the fun of it: Another bad plot
- Forcably made a pie chart with ggplot

```{r}
thirtysix<-ggplot(MOI_Determin, aes(x="", y=MOI_Determin$`Cell Density (36hrs)`, fill=MOI))+geom_bar(stat="identity", width=1 )+ coord_polar("y",start=0)

thirtysix

```


