---
title: "Hmk_04 template: Data frames and data wrangling"
format: gfm
editor: visual
---

Please read the entire [R for Data Science](https://r4ds.hadley.nz/data-transform) before you do this homework.

This homework relies on the `nycflights13` package, which contains several data frames, including `airlines`, `airports`, `flights`, `planes`, and `weather`. Loading `nycflights13` (`library(nycflights13)`) puts all of these data frames on the search path.

## Installing data packages

Remember that any package needs to be installed only once (per version of R), but needs to be loaded every time we start a new R session.

To install `nycflights13`, use `install.packages("nycflights13")` (with quotation marks). To load it, use `library(nycflights13)`.

```{r}
library(nycflights13)
library(tidyverse)

```

# Question 1: filtering

Make a plot of air time as a function of distance (air time on the y axis, distance on the x axis) for all flights that meet the following criteria:

-   originate from LaGuardia airport ("LGA")
-   departed on the 16th of the month
-   have a flight distance of less than 2000

```{r}
result<- flights |>
  filter(origin=="LGA")|>
  filter(day==16) |>
  filter(distance<2000)

plt<-ggplot(data=result, mapping = aes(distance, air_time))+geom_point()

suppressWarnings(print(plt))
```

# Question 2: dealing with NAs

Make a data frame of all of the rows of `flights` that have values for *both* `arr_time` and `dep_time` - that is, neither of those values are `NA`.

```{r}
resultQ2 <- flights |>
  filter(!is.na(arr_time)) |>
  filter(!is.na(dep_time))

```
Question/For me to think more about: How do I ask R to look for any NA in the final df to confirm it worked? I used print(is.na(resultQ2$arr_time)) and print(is.na(flights$arr_time)) for a TRUE/FALSE. 

## filtering NAs

`ggplot()` will automatically remove NA values from the plot, as you may have seen in question 1, but it emits a warning message about that. Of course you could silence the warning message using [chunk options](https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html), but how could you prevent them from appearing in the first place?

ANS: I can prevent the warning message by assigning the plot to a variable and using the suppressWarning( ) function when I display the plot.

```{r}
plt2<-ggplot(data=result, mapping = aes(distance, air_time))+geom_point()

suppressWarnings(print(plt2))
```

# Question 3: adding columns

Create a data frame of average flight speeds, based on `air_time` and `distance`. Make either a histogram or a density plot of the data. If you like, you may break the data out (e.g. by airline, or some other variable) in a way that you think makes sense.

```{r}
resultQ3 <- flights |>
  mutate(speed = distance / air_time) |>
  group_by(month) |>
  mutate(avg_flight_speed = mean(speed, na.rm=TRUE))

plt3 <- ggplot(data=resultQ3, mapping =aes(avg_flight_speed)) + geom_density()
print(plt3)

```
